Cosmological constraints from the redshift dependence of the volume effect: galaxy angular 2-point correlation function

### 2017-04-01 draft/draft2.1.tex
updates

\begin{abstract}
We propose to use the redshift dependence of the galaxy angular 2-point correlation function (2pCF)
to constrain cosmological parameters.
When an incorrect set of parameters is assumed for the coordinate transformation from redshift space to comoving space, 
it produces {\it redshift-dependent mis-scaling} in the constructed galaxy distribution.
As a result, we observe a redshift-dependent rescaling of the measured galaxy angular 2pCF.
Using mock galaxy catalogues (containing millions of galaxies at redshifts 0, 0.5, 1, 1.5, 2) from the Horizon Run 4 N-body simulation, we find that the redshift evolution of the angular 2pCF shape, as a function of scale,
is sensitive to the adopted cosmological parameters. 
% mock galaxy catalogues having hundreds of millions of galaxies at redshifts 0, 0.5, 1, 1.5, 2.
However, it is relatively insensitive to the gravitational growth of structure,
the galaxy bias, and the effect of redshift space distortions.
Analyzing the 2pCF shape on scales $5  h^{-1} {\rm Mpc} \leq r_\perp \leq 40 h^{-1} {\rm Mpc}$, 
we derive tight constraints on $\Omega_m$ and $w$ with 68.3\% CL uncertainties of 0.015 and 0.05.
%Our method is complementary to current works using 2pCF to constrain cosmological parameters.
The method could be applicable to future large scale structure surveys, especially photometric surveys such as DES, LSST, 
to derive tight cosmological constraints.
This work is a continuation of our previous works \citep{Li2014,Li2015,Li2016}
as a strategy to constrain cosmological parameters using redshift-invariant physical quantities;
they are all powerful in constrain cosmology and have a lot of merits,
e.g., are less affected by the systematic effects introduced by non-linear clustering or RSD,
does not require accurate modeling of galaxy clustering,
and are applicable on relative small clustering scales.
\end{abstract}

@@

\begin{figure*}
   \centering{
   \includegraphics[width=16cm]{fig4_a.eps}
   \includegraphics[width=16cm]{fig4_b.eps}
   }
   \caption{\label{fig_sys}
  Upper panels: $\omega_{r_\perp}$ (left) and $\hat{\omega}_{r_\perp}$ (middle) are shown for four different halo-mass-cuts, $3\times 10^{11},~1\times 10^{12},~4\times 10^{12},~\&~1\times 10^{13}~M_\odot$, 
  below which we remove from 2pCF calculation. 
  The right panel shows the difference in $\hat{\omega}_{r_\perp}$ between the $3\times 10^{11}~M_\odot$ mass-cut case and the other mass-cut cases. 
  Lower panels: $\omega_{r_\perp}$ (left), $\hat{\omega}_{r_\perp}$ (middle), and $\hat\omega_{r_\perp}(z=z_i)-\hat\omega_{r_\perp}(z=0)$ (right) calculated with the RSD effect included.
   }
\end{figure*}

In Sec. \ref{sec_2pCF_diffz} we showed that $\hat \omega_{r_\perp}$ measured from the
constant mass cut samples at different redshifts show good agreement.
The gravitational growth of structure will not have a large impact on $\hat \omega_{r_\perp}$ on scales $\gtrsim5 h^{-1}$ Mpc,
yet there are many other factors that may induce redshift evolution
including galaxy bias, RSD, redshift error, and 
the redshift evolution of galaxies properties 
such as mass, morphology, color, concentration. 
Here we test two ``major'' systematical effects; galaxy bias and the RSD effect.

@@

We vary the mass cut and check the effect of galaxy bias on $\hat \omega_{r_\perp}$.
The upper panels of Figure \ref{fig_sys} displays the 2pCF measured from galaxies within a $3175\times3175\times105$ $(h^{-1}\rm Mpc)^3$ volume, taken from the $z=0$ snapshot.
Four values of minimal mass limits, $3\times 10^{11} M_{\odot}$, $1\times 10^{12} M_{\odot}$, $4\times 10^{12} M_{\odot}$ and $1\times 10^{13} M_{\odot}$,
are imposed to create samples with different galaxy biases.
The measured angular 2pCF are displayed.

\begin{figure*}
   \centering{
   %\includegraphics[width=7cm]{Tpcf--contour.eps}
   %\includegraphics[width=16cm]{Tpcf--contour-others.eps}
   %\includegraphics[width=16cm,natwidth=52,natheight=40]{fig10.eps}
   \includegraphics[width=12cm]{fig9.eps}
   }
   \caption{\label{fig_contours}
   Likelihood contours (68.3\%, 95.4\%) in the $\Omega_m-w$ plane from our method, 
   based on the redshift evolution of the angular 2pCF shape $\hat\omega_{r_\perp}$.
   Utilizing the HR4 snapshots data of mock galaxies at five redshifts,
   the method can lead to very tight cosmological constraint.
   }
\end{figure*}

@@

The lower panel of Figure \ref{fig_sys} shows how it affects on the 2pCF.
Comparing this plot with Figure \ref{fig_diffz} one can clearly see the effect of RSD.
The amplitude of the measured 2pCF is enhanced by $\sim 10\%$ when considering the RSD effect 
and the slop of $\hat\omega_{r_\perp}$ is suppressed.
However, the redshift evolution of $\hat\omega_{r_\perp}$ remains small,  $\lesssim4\%$ at $0.5\leq z \leq 2$.
Therefore RSD should not significantly affect the application of our method, as long as its redshift evolution is small;
the small effect of RSD can be precisely modelled by simulations and corrected.
%If looking at the redshift evolution,
%the curves $\hat{r_\perp\omega} - \hat{r_\perp\omega}(z=0)$ displayed in the right panel,
%they changed a lot compared with the case of no RSD effect (the right panel of Figure \ref{fig_diffz}),
%but do not become much larger. %But RSD effect does not introduce large redshfit evolution, although the shape of  is

@@

@@

We constrain $\Omega_m$ and $w$ through a Bayesian analysis (\cite{Bayesian}; also see \cite{LB2002,Li2016}).
We assume that the likelihood takes the form
\begin{equation}\label{eq:like}
 \mathcal{L} \propto \exp\left[-\frac{\chi^2}{2}\right],
\end{equation}
and scan the parameter space in the $\Omega_m-w$ plane to obtain the 68.3\% and 95.4\% CL regions.
The result is displayed in Figure \ref{fig_contours}.

@@

>>>>

\begin{abstract}
{\bf Let us use 1) angular CF across the line-of-sight 2) $\xi_{\bot}$ rather than $\omega_{r_{\bot}}$}
We propose to use the redshift dependence of the galaxy angular 2-point correlation function (2pCF)
to constrain cosmological parameters.
When an incorrect set of parameters is assumed for the coordinate transformation from redshift space to comoving space, 
it produces {\it redshift-dependent mis-scaling} in the constructed galaxy distribution.
As a result, we observe a redshift-dependent rescaling of the measured galaxy angular 2pCF.
Using mock galaxy catalogues (containing millions of galaxies at redshifts 0, 0.5, 1, 1.5, 2) from the Horizon Run 4 N-body simulation, we find that the redshift evolution of the angular 2pCF shape, as a function of scale,
is sensitive to the adopted cosmological parameters. 
% mock galaxy catalogues having hundreds of millions of galaxies at redshifts 0, 0.5, 1, 1.5, 2.
However, it is relatively insensitive to the gravitational growth of structure,
the galaxy bias, and the effect of redshift space distortions.
Analyzing the 2pCF shape on scales $5  h^{-1} {\rm Mpc} \leq r_\perp \leq 40 h^{-1} {\rm Mpc}$, 
we derive tight constraints on $\Omega_m$ and $w$ with 68.3\% CL uncertainties of 0.015 and 0.05.
%Our method is complementary to current works using 2pCF to constrain cosmological parameters.
The method could be applicable to future large scale structure surveys, especially photometric surveys such as DES, LSST, 
to derive tight cosmological constraints.
This work is a continuation of our previous works \citep{Li2014,Li2015,Li2016}
as a strategy to constrain cosmological parameters using redshift-invariant physical quantities;
they are all powerful in constrain cosmology and have a lot of merits,
e.g., are less affected by the systematic effects introduced by non-linear clustering or RSD,
does not require accurate modeling of galaxy clustering,
and are applicable on relative small clustering scales.
\end{abstract}

@@

\begin{figure*}
   \centering{
   \includegraphics[width=16cm]{fig4_a.eps}
   \includegraphics[width=16cm]{fig4_b.eps}
   \includegraphics[width=16cm]{fig4_c.eps}
   \includegraphics[width=16cm]{fig4_d.eps}
   }
   \caption{\label{fig_sys}
  First row: $\omega_{r_\perp}$ (left), $\hat{\omega}_{r_\perp}$ (middle), and $\hat\omega_{r_\perp}(z=z_i)-\hat\omega_{r_\perp}(z=0)$ (right) calculated with the RSD effect included.
  Second row: $\omega_{r_\perp}$ (left) and $\hat{\omega}_{r_\perp}$ (middle) are shown for four different halo-mass-cuts, $3\times 10^{11},~1\times 10^{12},~4\times 10^{12},~\&~1\times 10^{13}~M_\odot$, 
  below which we remove from 2pCF calculation. 
  The right panel shows the difference in $\hat{\omega}_{r_\perp}$ between the $3\times 10^{11}~M_\odot$ mass-cut case and the other mass-cut cases. 
  Third row: The same as the second row panels, except that the RSD effect is considered.
  Fourth row: The redshift evolution in case of using {\it samples with constant number density} $\bar n=7.3\times 10^{-3} {\rm Mpc}^{-3}h^3$, 
  and the RSD effect is considered. } 
\end{figure*}

In Sec. \ref{sec_2pCF_diffz} we showed that $\hat \omega_{r_\perp}$ measured from the
constant mass cut samples at different redshifts show good agreement.
The gravitational growth of structure will not have a large impact on $\hat \omega_{r_\perp}$ on scales $\gtrsim5 h^{-1}$ Mpc,
yet there are many other factors that may induce redshift evolution
including galaxy bias, RSD, redshift error, and 
the redshift evolution of galaxies properties 
such as mass, morphology, color, concentration. 
Here we test two ``major'' systematical effects;  the RSD effect and galaxy bias.

@@

@@

The upper panel of Figure \ref{fig_sys} shows how it affects on the 2pCF.
Comparing this plot with Figure \ref{fig_diffz} one can clearly see the effect of RSD.
The amplitude of the measured 2pCF is enhanced by $\sim 10\%$ when considering the RSD effect 
and the slop of $\hat\omega_{r_\perp}$ is suppressed.
However, the redshift evolution of $\hat\omega_{r_\perp}$ remains small,  $\lesssim4\%$ at $0.5\leq z \leq 2$.
Therefore RSD should not significantly affect the application of our method, as long as its redshift evolution is small;
the small effect of RSD can be precisely modelled by simulations and corrected.
%If looking at the redshift evolution,
%the curves $\hat{r_\perp\omega} - \hat{r_\perp\omega}(z=0)$ displayed in the right panel,
%they changed a lot compared with the case of no RSD effect (the right panel of Figure \ref{fig_diffz}),
%but do not become much larger. %But RSD effect does not introduce large redshfit evolution, although the shape of  is

@@

\begin{figure*}
   \centering{
   %\includegraphics[width=7cm]{Tpcf--contour.eps}
   %\includegraphics[width=16cm]{Tpcf--contour-others.eps}
   %\includegraphics[width=16cm,natwidth=52,natheight=40]{fig10.eps}
   \includegraphics[width=8cm]{fig9_a.eps}
   \includegraphics[width=8cm]{fig9_b.eps}
   }
   \caption{\label{fig_contours}
   Left: Likelihood contours (68.3\%, 95.4\%) in the $\Omega_m-w$ plane from our method, 
   based on the redshift evolution of the angular 2pCF shape $\hat\omega_{r_\perp}$.
   Utilizing the HR4 snapshots data of mock galaxies at five redshifts,
   the method can lead to very tight cosmological constraint.
   Right: The constraints in case that one or two high redshift bins are not used. 
   }
\end{figure*}

We vary the mass cut and check the effect of galaxy bias on $\hat \omega_{r_\perp}$.
The second row of Figure \ref{fig_sys} displays the 2pCF measured from galaxies within a $3175\times3175\times105$ $(h^{-1}\rm Mpc)^3$ volume, taken from the $z=0$ snapshot.
Four values of minimal mass limits, $3\times 10^{11} M_{\odot}$, $1\times 10^{12} M_{\odot}$, $4\times 10^{12} M_{\odot}$ and $1\times 10^{13} M_{\odot}$,
are imposed to create samples with different galaxy biases.
The measured angular 2pCF are displayed.

% {\bf 
The combined effect of galaxy bias and RSD results in larger systematics.
Galaxies that are more biased predominantly reside within nonlinear, over-dense regions and thus have larger larger peculiar velocities.
The third row of Figure \ref{fig_sys} shows that, 
the difference of $\hat \omega_{r_\perp}$ among different mass cut samples reaches $1.5\%,\ 4\%$ and $8\%$ in the case when including the RSD effect.

Finally, the fourth row of the Figure displays the 
redshift evolution in case of using samples with constant number density $\bar n=7.3\times 10^{-3} {\rm Mpc}^{-3}h^3$, 
with the RSD effect considered. %(the error bar is a bit larger since for an illustration purpose we only do the computation for 1/30 of the whole sample).
We see the redshfit evolution of $\hat \omega_{r_\perp}$ is as small as $\lesssim3\%$.

In summary, the effect of RSD and galaxy bias on the shape of galaxy angular correlation function is, not large, but it is still non-negligible.
When dealing with real observational data, one should construct simulation mocks in which these effects are reliably modelled, 
so that the systematics are corrected and will not lead to significant bias in the derived cosmological constraints.

@@

We constrain $\Omega_m$ and $w$ through a Bayesian analysis (\cite{Bayesian}; also see \cite{LB2002,Li2016}).
We assume that the likelihood takes the form
\begin{equation}\label{eq:like}
 \mathcal{L} \propto \exp\left[-\frac{\chi^2}{2}\right],
\end{equation}
and scan the parameter space in the $\Omega_m-w$ plane to obtain the 68.3\% and 95.4\% CL regions.
The result is displayed in the left panel Figure \ref{fig_contours}.

@@

In case that we do not have high redshfit galaxies, the constraint becomes weaker.
Figure \ref{fig_contours} displays that, 
in case that we only use the galaxies at $z=$0, 0.5 and 1, 
the 2$\sigma$ uncertainties of $\delta\Omega_m\approx0.01,\\ \delta w\approx0.1$ (the other parameter fixed at its best-fit value).
The area of contour is further enlarged by $\approx 50\%$ if we only use galaxies at $z=$0 and 0.5.
